\documentclass[a4paper, 12pt]{paper}
\input{preamble}

\begin{document}
{\noindent
	v 1.1 \\
  \\
{\Huge\textbf{A Contraction-free System for STL}}
}
\\
\part*{Definitions}
\setcounter{section}{-1}
\section{Notation} In the following, a formula is a free construction of operators $\bot$, $\top$, $\nabla$, $\vee$, $\wedge$ and $\rightarrow$, which have arities $0$, $0$, $1$, $2$, $2$ and $2$, respectively, over a countable set of atoms. $\Gamma$, $\Sigma$ and $\Pi$ are names for finite multi-sets of formulas, $\Delta$ for a sub-singleton of some formula, $A$, $B$ and $C$ for formulas, $p$ for atoms and $n$, $l$, $r$ and $k$ for natural numbers. We denote by $P$ the set of all atoms, and by $P(A)$ the set of those that occur in the formula $A$.

We will write $A$ for the singleton $\{A\}$ where ever it is inferable from the context.
``$,$'' is the multi-set union. We will also use the following short-hands.
\begin{flushleft}
  $ A^1 = \{ A \} $ \\
  $ A^{n+1} = A^n, A $ \\
  $ \Gamma^1 = \Gamma $ \\
  $ \Gamma^{n+1} = \Gamma^n, \Gamma $ \\
  $ \nabla^1 A = \nabla A $ \\
  $ \nabla^{n+1} A = \nabla (\nabla^n A) $ \\
  $ \nabla^n \Gamma = \{ \nabla^n A \mid A \in \Gamma \} $ \\
  $ P(\Gamma) = \bigcup_{A \in \Gamma} P(A) $
\end{flushleft}


A sequent $\Gamma \Rightarrow \Delta$ is a binary relation between $\Gamma$, a multi-set of formulas called \emph{the antecedent} or \emph{the left-side}, and $\Delta$, a sub-singleton of some formula called \emph{the succedent} or \emph{the right-side}.


We will construct trees, called \emph{proof-trees}, using a set of recursive rules, called a \emph{deductive system}, or simply \emph{system}.

A rule $R$ is expressed as a relation between a set of $n$ sequents $\{ \Gamma_i \Rightarrow \Delta_i \mid 1 \leq i \leq n \}$ called \emph{premises} and a sequent $\Gamma \Rightarrow \Delta$ called \emph{conclusion}, and is written as follows.
\begin{prooftree}
  \AXC{$\Gamma_1 \Rightarrow \Delta_1$}
  \AXC{$\dots$}
  \AXC{$\Gamma_n \Rightarrow \Delta_n$}
  \RightLabel{$R$}
  \TIC{$\Gamma \Rightarrow \Delta$}
\end{prooftree}

We will designate a specific formula in the conclusion of some rules, called \emph{the principal formula} of that rule, which will be indicated by writing it \uwave{underlined}.

A proof-tree in a system is made up of instances of rules of that system. So in a system with the rule $R$, we can construct a proof-tree with root $\Gamma \Rightarrow \Delta$, if we have already constructed $n$ proof-trees with roots $\Gamma_i \Rightarrow \Delta_i$ for all $1 \leq i \leq n$.



When a proof-tree, with $\Gamma \Rightarrow \Delta$ as root, is constructed using a system $S$, we say that $S$ proves $\Gamma \Rightarrow \Delta$, and write it as $S \vdash \Gamma \Rightarrow \Delta$.



We will name proof-trees by $\D$, $\D'$ and so on, unless otherwise stated. We will name subtrees of $\D$ by $\D_0$, $\D_1$ and so on. We will frequently use induction to prove existence of some proof-tree, assuming existence of some other proof-tree $\D$. In such situations we will use notation $IH(\D)$ to denote the proof-tree that we will get from induction hypothesis.

We will write $S \vdash_h \Gamma \Rightarrow \Delta$ to indicate the existence of a proof-tree of height $h$ for $\Gamma \Rightarrow \Delta$ in system $S$. We will also write $h(\D)$ for the height of a proof-tree $\D$.

We write $S \vdash^r A$ when we want to indicate that a system $S$ proves a formula $A$, with a proof-tree of rank $r$ (defined below).


\section{cut} \quad \\

\begin{center}
  \begin{prooftree}
    \AXC{$ \Gamma \Rightarrow A$}
    \AXC{$\Sigma, A \Rightarrow \Delta$}
    \RightLabel{$cut$}
    \BIC{$\Gamma, \Sigma \Rightarrow \Delta$}
  \end{prooftree}
\end{center}

\begin{center}
  \begin{prooftree}
    \AXC{$ \Gamma \Rightarrow A$}
    \AXC{$\Sigma, \nabla^n A \Rightarrow \Delta$}
    \RightLabel{$\nabla cut$}
    \BIC{$\nabla^n \Gamma, \Sigma \Rightarrow \Delta$}
  \end{prooftree}
\end{center}

Notice that $\nabla cut$ generalizes $cut$.

$A$ is called \emph{the cut-formula} of a $cut$ or $\nabla cut$ instance.

\emph{Rank} of a formula $A$ is defined as
\[ \rho(A) = \begin{cases}
1 & \quad ; A \in P \cup \{ \bot, \top \} \\
\rho(B) & \quad ; A = \nabla B \\
max(\rho(B), \rho(C)) + 1 & \quad ; A = B \circ C \quad (\circ \in \{ \land, \lor, \rightarrow \})
\end{cases} \]
Notice that $\nabla$ does not increase the rank of a formula.

We also define rank for rule instances and proof-trees as follows. Rank of an instance of the $\nabla cut$ rule with cut-formula $A$ is defined to be the rank of $A$. Rank of any other rule instance is $0$.
Rank of a proof-tree $\D$, written as $\rho(\D)$, is the maximum rank of all of its rule instances.


\section{ST} \quad \\

\input{dfn/st}

By $\st \vdash \Gamma \Rightarrow \Delta$ we mean that the sequent $\Gamma \Rightarrow \Delta$ provable in this system.

By $\st^+ \vdash \Gamma \Rightarrow \Delta$ we mean that the sequent $\Gamma \Rightarrow \Delta$ provable in the above system, augmented with the $cut$ rule. 
We denote the \emph{spacetime logic} by $\stl$, which is the set of all sequents provable in $\st^+$.



\section{ST3} \quad \\

\input{dfn/stt}


Notice that there could be more than one principal-formula in $Lw$.

By $\stt \vdash \Gamma \Rightarrow \Delta$ we mean that the sequent $\Gamma \Rightarrow \Delta$ provable in this system.

By $\stt^+ \vdash \Gamma \Rightarrow \Delta$ we mean that the sequent $\Gamma \Rightarrow \Delta$ provable in the above system, augmented with the $cut$ rule.



\section{GST3} \quad \\

\input{dfn/gstt}

Notice that all $\gstt$ rules generalize $\stt$ rules.

Also notice that there could be more than one principal-formula in $Lw$.

By $\gstt \vdash \Gamma \Rightarrow \Delta$ we mean that the sequent $\Gamma \Rightarrow \Delta$ provable in this system.

By $\gstt^+ \vdash \Gamma \Rightarrow \Delta$ we mean that the sequent $\Gamma \Rightarrow \Delta$ provable in the above system, augmented with the $\nabla cut$ rule.

We denote the logic of $\gstt$, which is the set of all sequents provable in $\gstt$, by $\gsttl$.


\section{Conventions} We will refer to $Id$, $L \bot$, $R \top$ as \emph{axioms}, $L \wedge$, $L \wedge_1$, $L \wedge_2$, $L \vee$, $R \vee_1$, $R \vee_2$, $L \rightarrow$, $R \rightarrow$ as \emph{logical rules}, $Lw$, $Rw$ and $Lc$ as \emph{structural rules} and $N$ as \emph{the modal rule}.


\pagebreak

\part*{Theorems}
\setcounter{section}{0}


\section{Theorem}\label{thm:stt-id-form} $(Id)$ $\stt \vdash \Gamma, A \Rightarrow A$. \hyperref[pr:stt-id-form]{\emph{Proof.}}


\section{Lemma}\label{lem:stt-inversion}
\begin{enumerate}
  \item If $\stt \vdash_h \Gamma, \nabla^n (A \wedge B) \Rightarrow \Delta$ then $\stt \vdash_h \Gamma, \nabla^n A, \nabla^n B \Rightarrow \Delta$.
  \item If $\stt \vdash_h \Gamma, \nabla^n (A \vee B) \Rightarrow \Delta$ then $\stt \vdash_h \Gamma, \nabla^n A \Rightarrow \Delta$ and $\stt \vdash_h \Gamma, \nabla^n B \Rightarrow \Delta$.
  \item If $\stt \vdash_h \Gamma, \nabla^n B, \nabla^{n+1} (A \rightarrow B) \Rightarrow \Delta$ then $\stt \vdash_h \Gamma, \nabla^n B^2 \Rightarrow \Delta$.
  \item If $\stt \vdash_h \Gamma \Rightarrow A \wedge B$ then $\stt \vdash_h \Gamma \Rightarrow A$ and $\stt \vdash_h \Gamma \Rightarrow B$.
\end{enumerate}
\hyperref[pr:stt-inversion]{\emph{Proof.}}


\section{Theorem}\label{thm:stt-lc-elim} $(Lc)$ If $\stt \vdash_h \Gamma, A^n \Rightarrow \Delta$ then $\stt \vdash_h \Gamma, A \Rightarrow \Delta$.
\hyperref[pr:stt-lc-elim]{\emph{Proof.}}



\section{Theorem}\label{thm:stt-eq-st} $\stt \vdash \Gamma \Rightarrow \Delta$ iff $\st \vdash \Gamma \Rightarrow \Delta$.
\hyperref[pr:stt-eq-st]{\emph{Proof.}}



\section{Theorem}\label{thm:sttp-eq-stp} $\stt^+ \vdash \Gamma \Rightarrow \Delta$ iff $\st^+ \vdash \Gamma \Rightarrow \Delta$.
\hyperref[pr:sttp-eq-stp]{\emph{Proof.}}



\section{Lemma}\label{lem:l-nabla-dist}
\begin{enumerate}
  \item $\stt^+ \vdash \nabla^n (A \vee B) \Rightarrow \nabla^n A \vee \nabla^n B$.

  \item $\stt^+ \vdash \nabla^n (A \wedge B) \Rightarrow \nabla^n A \wedge \nabla^n B$. 

  \item $\stt^+ \vdash \nabla^n (A \rightarrow B) \Rightarrow \nabla^n A \rightarrow \nabla^n B$.
\end{enumerate}
\hyperref[pr:l-nabla-dist]{\emph{Proof.}}



\section{Theorem}\label{thm:sttp-eq-gsttp} $\stt^+ \vdash \Gamma \Rightarrow \Delta$ iff $\gstt^+ \vdash \Gamma \Rightarrow \Delta$. \hyperref[pr:sttp-eq-gsttp]{\emph{Proof.}}



\section{Theorem}\label{thm:gstt-cut-reduction}
If $\gstt \vdash^{r_0} \Gamma \Rightarrow A$ and $\gstt \vdash^{r_1} \Sigma , \nabla^n A \Rightarrow \Delta$, where $r_0, r_1 < \rho(A)$, then $\gstt \vdash^{r_2} \nabla^n \Gamma, \Sigma \Rightarrow \Delta$ and $r_2 < \rho(A)$.\hyperref[pr:gstt-cut-reduction]{\emph{Proof.}}



\section{Theorem}\label{thm:gstt-eq-gsttp} $\gstt \vdash \Gamma \Rightarrow \Delta$ iff $\gstt^+ \vdash \Gamma \Rightarrow \Delta$. \hyperref[pr:gstt-eq-gsttp]{\emph{Proof.}}



\section{Corollary}\label{thm:gstt-eq-stp} $\gstt \vdash \Gamma \Rightarrow \Delta$ iff $\st^+ \vdash \Gamma \Rightarrow \Delta$. \hyperref[pr:gstt-eq-stp]{\emph{Proof.}}



\section{Corollary} $\stl = \gsttl$, thus, we have a cut-free system for $\stl$.


\pagebreak

\part*{Proofs}
\setcounter{section}{0}

\section{Proof of \ref{thm:stt-id-form}}\label{pr:stt-id-form}
\input{thm/stt-id-form}


\section{Proof of \ref{lem:stt-inversion}}\label{pr:stt-inversion} \input{thm/l-lconj-inv}



\section{Proof of \ref{thm:stt-lc-elim}}\label{pr:stt-lc-elim}
By induction on the proof for $\Gamma, A^{n+1} \Rightarrow \Delta$, which we call $\D$ and ends with rule $R$.
Cases for axioms are trivial. All cases where none of occurrences of $A$ are principal are handled by commuting $R$ with $IH(D_i)$. Now suppose $R$ is either of the following and $A$ is principal in $R$.
\begin{itemize}
  \item[$(L \wedge)$] Let $A = A \wedge B$.
  \begin{prooftree}
    \AXC{$\D_0$} \noLine
    \UIC{$\Gamma, A, B, (A \wedge B)^n \Rightarrow \Delta$}
    \RightLabel{$L \wedge$}
    \UIC{$\Gamma, (A \wedge B)^{n+1} \Rightarrow \Delta$}		
  \end{prooftree}
  Then by Lemma \ref{lem:stt-inversion} we have $\Gamma, A^{n+1}, B^{n+1} \Rightarrow \Delta$. Using induction hypothesis twice, we would have $\Gamma, A, B \Rightarrow \Delta$. $L \wedge$ will result in the desired sequent.

  \item[$(L \vee)$] Let $A = A \vee B$.
  \begin{prooftree}
    \AXC{$\D_0$} \noLine
    \UIC{$ \Gamma, A, (A \vee B)^n \Rightarrow \Delta$}
    \AXC{$\D_1$} \noLine
    \UIC{$\Gamma, B, (A \vee B)^n \Rightarrow \Delta$}
    \RightLabel{$L \vee$}
    \BIC{$ \Gamma, (A \vee B)^{n+1} \Rightarrow \Delta$}		
  \end{prooftree}
  By Lemma \ref{lem:stt-inversion} for $\D_0$ and $\D_1$ we have $\Gamma, A^{n+1} \Rightarrow \Delta$ and $\Gamma, B^{n+1} \Rightarrow \Delta$. By induction hypothesis we have $\Gamma, A \Rightarrow \Delta$ and $\Gamma, B \Rightarrow \Delta$. By $L \vee$ we have the desired sequent.

  \item[$(L \rightarrow)$] Let $A = \nabla (A \rightarrow B)$.
  \begin{prooftree}
    \AXC{$\D_0$} \noLine
    \UIC{$\Gamma, (\nabla (A \rightarrow B))^{n+1} \Rightarrow A$}
    \AXC{$\D_1$} \noLine
    \UIC{$\Gamma, B, (\nabla (A \rightarrow B))^n \Rightarrow \Delta$}
    \RightLabel{$L \rightarrow$}
    \BIC{$\Gamma, (\nabla (A \rightarrow B))^{n+1} \Rightarrow \Delta$}
  \end{prooftree}
  By Lemma \ref{lem:stt-inversion} for $\D_1$ we have $\Gamma, B^{n+1} \Rightarrow \Delta$. By induction hypothesis we have $\Gamma, \nabla (A \rightarrow B) \Rightarrow A$ and $\Gamma, B$. By $L \rightarrow$ we have the desired sequent.
\end{itemize}



\section{Proof of \ref{thm:stt-eq-st}}\label{pr:stt-eq-st}\quad \\
($\Rightarrow$) All the rules in $\stt$ are also in $\st$, except for $L \wedge$ and $L \rightarrow$, which can be easily imitated in $\st$ using the same logical rules and structural rules in $\st$. \\
($\Leftarrow$) All the rules in $\st$ are also in $\stt$, except for $Lc$, which can be imitated in $\stt$ using Theorem \ref{thm:stt-lc-elim}.



\section{Proof of \ref{thm:sttp-eq-stp}}\label{pr:sttp-eq-stp}\quad \\
Similar to \ref{pr:stt-eq-st}.\\
($\Rightarrow$) By induction on the structure of $\D$, the proof-tree for $\Gamma \Rightarrow \Delta$ in $\stt^+$, which ends with rule $R$.
For the cases where $R$ is also a rule in $\st^+$, just apply $R$ on the proof-trees that we get from induction hypothesis for the sub-tree(s) of $\D$. If $R$ is $L \wedge$, we must use both $L \wedge_1$ and $L \wedge_2$, and $Lc$ in $\st^+$. If $R$ is $L \rightarrow$, we must use $Lw$ to prepare the context before using $L \rightarrow$ in $\st^+$.\\
($\Leftarrow$) Similar to the other direction, except that in the case for $Lc$ we must use Theorem \ref{thm:stt-lc-elim}.


\section{Proof of \ref{lem:l-nabla-dist}}\label{pr:l-nabla-dist}\quad \\
\begin{enumerate}
  \item \quad \input{thm/nabla-disj}
  \item \quad \input{thm/nabla-conj}
  \item \quad \input{thm/nabla-impl}
\end{enumerate}



\section{Proof of \ref{thm:sttp-eq-gsttp}}\label{pr:sttp-eq-gsttp}\quad \\
($\Rightarrow$) Easily follows from the fact that all rules of $\stt^+$ are just instances of $\gstt^+$'s rules.\\
($\Leftarrow$) We will use induction on the proof-tree $\D$ for $\Gamma \Rightarrow \Delta$ in $\gstt^+$, which ends with rule $R$.

The cases for Axioms are trivial.
For the rules that are common between two systems, just apply $R$ on the proof-tree(s) from the induction hypothesis.
In the cases for $L \wedge$, $L \vee$ and $L \rightarrow$, do the same, and also $\nabla cut$ the sequents proved in Lemma \ref{lem:l-nabla-dist} into the resulting sequent. Finally, if $R$ is $\nabla cut$, just apply $N$ $n$ times before using $cut$ in $\stt^+$.



\section{Proof of \ref{thm:gstt-cut-reduction}}\label{pr:gstt-cut-reduction}
\input{thm/gstt-cut-reduction}



\section{Proof of \ref{thm:gstt-eq-gsttp}}\label{pr:gstt-eq-gsttp}
\input{thm/gstt-cut-elim}



\section{Proof of \ref{thm:gstt-eq-stp}}\label{pr:gstt-eq-stp}
Trivial.



\pagebreak

\part*{\emph{Notes}}

\begin{itemize}
  \item We can't eliminate weakening rules from $\stt$, because of $N$.

  \item $L \wedge$ in $\stt$ gives us (partial) normality.
  
  \item We need Inversion Lemmas and Contraction Admissibility Theorem to preserve length, because we will use induction more than one on the case for $L \wedge$ in Contraction Admissibility Theorem.
  
  \item We need $Lw$ in $\stt$ to add more than one formula to the left-side, at once, since we will have to do so in $L \wedge$ Admissibility Lemma, in the case where the $A \wedge B$ is principal in $Lw$.
  
  \item We can't prove inverse of $L \rightarrow$, because
  \begin{prooftree}
    \RightLabel{$Id$}
    \AXC{$p \Rightarrow p$}
    \RightLabel{$Id$}
    \AXC{$q \Rightarrow q$}
    \RightLabel{$L \rightarrow$}
    \BIC{$\nabla (p \rightarrow q), p \Rightarrow q$}
    \RightLabel{$R \rightarrow$}
    \UIC{$p \rightarrow q \Rightarrow p \rightarrow q$}
    \RightLabel{$N$}
    \UIC{$\nabla (p \rightarrow q) \Rightarrow \nabla (p \rightarrow q)$}
  \end{prooftree}
  but $q \Rightarrow \nabla (p \rightarrow q)$ needs (implies?) $Fa$.

  \item Inversion Lemmas for the left-rules are generalized with arbitrary number of $\nabla$s, because in the case for $R \rightarrow$ we would have a $\nabla$ added to the left-side of the premise.
  
  \item Not sure if we still have to eliminate generalized contraction.
  
\end{itemize}

\end{document}
